pegax
=====

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE
)
```

[![Build Status](https://travis-ci.org/ropenscilabs/pegax.svg?branch=master)](https://travis-ci.org/ropenscilabs/pegax)
[![codecov.io](https://codecov.io/github/ropenscilabs/pegax/coverage.svg?branch=master)](https://codecov.io/github/ropenscilabs/pegax?branch=master)


Taxonomy PEG (Parsing Expression Grammar) rules

## Example rules

Capture any letter

```
struct name
  : plus< alpha >
{};
```

Capture any digit

```
struct numbers
  : plus< digit >
{};
```

## Grammar

Rules are combined to form a grammar, 

e.g., string must match `name`, then have one comma, then one space, 
then match `numbers`.

```
struct grammar
  : must< name, one< ',' >, space, numbers, eof >
{};
```

Which is then applied to parsing user input strings

## Install

```{r eval=FALSE}
devtools::install_github("ropenscilabs/pegax")
```

```{r}
library("pegax")
```


## Parse authority

Author name

```{r}
pgx_authority_names("Linnaeus, 1758")
```

Author year

```{r}
pgx_authority_years("Linnaeus, 1758")
```

Moar

```{r}
x <- c('Smith, 1456', 'Foobarcheesestuff, 1001', 'Smith, 1987', "Linnaeus 2014", "Linnaeus 2014?", "Page 1234    ", "     Jones, 1789     ")
pgx_authority_names(x)
pgx_authority_years(x)
```

Even Moar

```{r}
library("charlatan")
per <- charlatan::PersonProvider$new()
date <- charlatan::DateTimeProvider$new()
x <- replicate(10^4, paste(
    sample(per$person$last_names, 1),
    date$year(),
    sep = ", "
  )
)
length(x)

# years
invisible(pgx_authority_years(x))
system.time(pgx_authority_years(x))
## vs. regex
### write me

# names
invisible(pgx_authority_names(x))
system.time(pgx_authority_names(x))
## vs. regex
### write me
```

## Parse ranks

```{r}
pgx_ranks("Fagus sylvatica subsp. orientalis")
```

Many at cone

```{r}
pgx_ranks(c("Helianthus annuus var. annuus", 
   "Helianthus annuus ssp. annuus", "Caulerpa cupressoides forma nuda"))
```

## Parse complete names

```{r}
pgx_parse("Tridentella tangeroae Bruce, 198?")
```

## R DSL to make custom PEGS for taxonomy

not made yet - but we can dream

```
pegax({
  plus(alpha)
  one(,)
  numbers(,)
  plus(-)
  space
  plus(????)
  plus(:cat2:)
})
```

## Meta

* citing this package: run `citation(package = "pegax")`
* Please note that this project is released with a [Contributor Code of Conduct](CODE_OF_CONDUCT.md).
  By participating in this project you agree to abide by its terms.
  